---
output:
  word_document: default
  html_document: default
---

##Problem1
###(a) 
Model the distribution using a mixture of K Gaussians with 
K = 3, 4, 5. Use BIC and AIC to find the best model.

```{r}
library(ggplot2)
library(mvtnorm)
library(bayesm)
library("mclust")
set.seed(123)
data=read.table("/Users/jing/Desktop/235/hw1/clustering data.txt",sep=',',header=T)
x1=data$X1
x2=data$X2
#jpeg('../figures/softKmeansData.jpeg')
p <- ggplot(data, aes(x1, x2)) + xlim(c(-10, 15)) + ylim(-16, 8)
p + geom_point()
#dev.off()
mod3 <- densityMclust(data,3)
mod4 <- densityMclust(data,4)
mod5 <- densityMclust(data,5)
AIC(mod3)
BIC(mod3)
AIC(mod4)
BIC(mod4)
AIC(mod5)
BIC(mod5)
```
From the plot of the original data,we could find the data is clustering in 4 groups.
We could find that when K=4,the AIC and BIC achieve the minimum values.

###(b) 
Use the optimum K from part (a) to cluster data using K-Means clustering.
```{r}
# K-means clustering with K=4
X=data
N <- nrow(X)
K = 4
set.seed(123)
# Initializing the centroids 
Cent <- cbind(runif(K, min(X[, 1]), max(X[, 1])), runif(K, min(X[, 2]), max(X[, 2])) )
D <- matrix(NA, N, K)
# run the algorithm for 500 iterations; in general, we should stop when the assignments don't change anymore.
for(i in 1:500){
  # Calculating the distance of each data point from the K centroids.   
  for (k in 1:K){
    D[, k] <- rowSums( (sweep(X, 2, Cent[k, ]))^2)
  }
  
  # Assigning the data points to the closest centroid  
  clus <- apply(D, 1, which.min)
  
  # Finding the new centroids   
  Cent <- by(X, INDICES=clus, FUN=colMeans)  
  Cent <- do.call(rbind, Cent)  
  
}  
data <- data.frame(x1 = X[, 1], x2 = X[, 2], clus = factor(clus))

#jpeg('../figures/kmeansResults.jpeg')
p <- ggplot(data, aes(x1, x2))
p + geom_point(aes(color=clus, shape=clus)) 
```
We could find that the result of K-Means clustering is consistent with the fact.



###(c) 
Use a hierarchical clustering method to cluster the data. Then use the optimum K from
part (a) to divide the observations into K groups.
```{r}
hc <- hclust(dist(X), "single")
dend <- as.dendrogram(hc)

plot(dend, leaflab = "none")
abline(h=4, lty=2, lwd=2)
clus <- cutree(hc, k=4)

data <- data.frame(x1 = X[, 1], x2 = X[, 2], clus = factor(clus))

#jpeg('../figures/kmeansResults2.jpeg')
p <- ggplot(data, aes(x1, x2))
p + geom_point(aes(color=clus, shape=clus)) 

```
We use single linkage clustering in hierarchical clustering, the result is consistent with the fact.



##Problem 2
The dataset communitiesCrimeRaw2 includes 19 measurements on 302 communities across the US. 

We first find the averages of these measurements at the state level so each row corresponds to one state.We have 44 states across the US.

Then reduce the dimension using PCA and factor analysis. 
```{r}
CR=read.table("/Users/jing/Desktop/235/communitiesCrimeRaw2.csv",sep=',',header=T)
CR=as.data.frame(CR)
means=aggregate(CR[,4:21],list(CR$state),mean)
CR.s <- scale(means[,2:19])

#PCA
pca.res = princomp(na.omit(CR.s))
summary(pca.res)
#loadings(pca.res) # small values are not printed
z <- pca.res$scores # scores 
pdf('pca_states.pdf')
biplot(pca.res)
dev.off()

#scree plot
plot(pca.res)
df <- data.frame(PC = 1:length((pca.res$sdev)), Variance = (pca.res$sdev)^2)
pdf('scree_states.pdf')
qplot(PC, Variance, data= df, geom='line')
dev.off()
```
From the scree plot, we could find the first 3 columns of z could explain the most part of the variance already,when achieve the first 10 columns of z, the variance is approximately 0.
```{r}
#Factor Analysis
fa.res <- factanal(na.omit(CR.s), 2, scores='regression')
print(fa.res)

pdf('fa_states.pdf')
biplot(fa.res$scores[, 1:2], fa.res$loadings[,1:2],xlim=c(-3, 3), ylim=c(-1, 3))
dev.off()
```
The 63% of the variance explained by three factors is enough.

##Problem 3
The dataset “Forest Fires"includes 512 data with 13 measurements that could predict the burned area of the forest.


We Use caret package in R to compare the following models via a 5-fold cross validation.
```{r}
library(caret)
ff=read.table("/Users/jing/Desktop/235/hw1/forestfires.csv",sep=',',header=T)
ff1=ff[,-c(3,4)]
ff1<-cbind(ff1,as.numeric(ff$month),as.numeric(ff$day))
ff1<-scale(ff1)
# Define train control for k fold cross validation
train_control <- trainControl(method="cv", number=5)
# Fit linear Model
set.seed(12345)
model1 <- train(area~., data=ff1, trControl=train_control, method="lm")
# Summarise Results
print(model1)

# Fit pcr Model
set.seed(12345)
model2 <- train(area~., data=ff1, trControl=train_control, method="pcr")
# Summarise Results
print(model2)

# Fit pls Model
set.seed(12345)
model3 <- train(area~., data=ff1, trControl=train_control, method="pls")
# Summarise Results
print(model3)

# Fit ridege Model
set.seed(12345)
model4 <- train(area~., data=ff1, trControl=train_control, method="ridge")
# Summarise Results
print(model4)

# Fit lasso Model
set.seed(12345)
model5 <- train(area~., data=ff1, trControl=train_control, method="lasso")
# Summarise Results
print(model5)
```